---
title: "Analyzing local loyalty"
output: pdf_document
---

This organized code series, accompanied by section headings, provides an in-depth exploration of the methods employed for analyzing local loyalty. The code reads the data once the necessary program packages are loaded. Please note that, regrettably, the database is not publicly accessible. The data was collected through a questionnaire where respondents provided ratings using a Likert scale.

Structural Equation Modeling (SEM) was applied to assess the relationships between demographic variables and local loyalty. However, it's worth mentioning that SEM can be adapted for various correlation analyses.

The Chi-squared (Chi2) test is a valuable tool for visualizing and quantifying associations between categorical variables. In this code series, the mosaic() function is used to perform an extensive statistical analysis of cross-tabulated data.

LCA cluster analysis was utilized to define loyalty groups, treating them as categorical variables for subsequent analyses.

To determine loyalty group membership, machine learning techniques were employed, specifically a random forest procedure. A classification task was defined, with explanatory variables serving as predictors and loyalty groups, as defined by LCA, acting as outcome variables. The Random Forest model identifies key predictors, which are also detailed in the related article.

For response category-level analyses, a correlation network was generated using the ARM package. This approach allows for a deeper exploration of the data's intricacies.

#Libraries

```{r}
#Data wrangling
library(dplyr)
library(reshape2)

#Chi2
library(vcd)

#SEM
library(seminr)
library(fastDummies)

#LCA
library(poLCA)

#Radom Forest
library(randomForest)

#ARM
library(arules)
library(arulesViz)

#plots 
library(visNetwork)
library(ggplot2)
library(extrafont)
library(stringr)
```

#Read data

```{r}
resp <- read.csv("resp.csv", sep = ";")
```

#SEM (Structural Equation Modelling)

##Age

```{r}
#what to test
dat <- resp %>% dplyr::select("Q1_1","Q6_1","Q6_2","Q29_4", #local loyalty
                              "gen")
colnames(dat) <- c("loy1","loy2","loy3","loy4","gen")
```

```{r}
table(dat$loy1)
table(dat$loy2)
table(dat$loy3)
table(dat$loy4)
table(dat$gen)
```

```{r}
dat <- dat %>% mutate(loy1=dplyr::recode(loy1, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>% 
  mutate(loy2=dplyr::recode(loy2, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy3=dplyr::recode(loy3, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy4=dplyr::recode(loy4, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(gen=dplyr::recode(gen, '1'='1', '2'='1', '3'='2', '4'='3', '5'='4')) #merging of older generations

dat[] <- sapply(dat, as.numeric)
dat <- dat[complete.cases(dat),]
```

```{r}
dat <- dummy_cols(dat, select_columns = c("loy1","loy2","loy3","loy4","gen"))
head(dat)
```

```{r}
# Create measurement model
simple_mm <- constructs(
  composite("Loyalty", multi_items("loy", 1:4)),
  composite("X", single_item("gen_2")),
  composite("Y", single_item("gen_3")),
  composite("Z", single_item("gen_4")))

# Create structural model
simple_sm <- relationships(
  paths(from = c("X", "Y", "Z"), to = "Loyalty"))

# Estimate the model
simple_model <- estimate_pls(data = dat,
 measurement_model = simple_mm,
 structural_model = simple_sm,
 missing = mean_replacement)

# Summarize the model results
summary_simple <- summary(simple_model)
summary_simple
```

```{r}
# Bootstrap the model on the PLS Estimated Model
boot_model <- bootstrap_model(
  seminr_model = simple_model,
  nboot = 5000,
  cores = parallel::detectCores(),
  seed = 123)

# alpha sets the specified level for significance, i.e. 0.05
summary_boot <- summary(boot_model, alpha = 0.05)

# Inspect the bootstrapped structural paths
summary_boot$bootstrapped_paths
```

##Financial circumastances

```{r}
#what to test
dat <- resp %>% dplyr::select("Q1_1","Q6_1","Q6_2","Q29_4", #local loyalty
                              "Q59")
colnames(dat) <- c("loy1","loy2","loy3","loy4","fin")
```

```{r}
table(dat$fin)
```

```{r}
dat <- dat %>% mutate(loy1=dplyr::recode(loy1, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>% 
  mutate(loy2=dplyr::recode(loy2, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy3=dplyr::recode(loy3, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy4=dplyr::recode(loy4, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
 # mutate(fin=dplyr::recode(fin, '1'='1', '2'='2', '3'='3')) %>% 
  mutate_at('fin', ~replace_na(.,-99))

dat[] <- sapply(dat, as.numeric)
dat <- dat[complete.cases(dat[ , c('loy1', 'loy2', 'loy3', 'loy4')]),]
```

```{r}
dat <- dummy_cols(dat, select_columns = c("loy1","loy2","loy3","loy4","fin"))
head(dat)
```

```{r}
# Create measurement model
simple_mm <- constructs(
  composite("Loyalty", multi_items("loy", 1:4)),
  composite("Poor", single_item("fin_1")),
  composite("Average", single_item("fin_2")))

# Create structural model
simple_sm <- relationships(
  paths(from = c("Poor", "Average"), to = "Loyalty"))

# Estimate the model
simple_model <- estimate_pls(data = dat,
 measurement_model = simple_mm,
 structural_model = simple_sm,
 missing = mean_replacement,
 missing_value = "-99")

# Summarize the model results
summary_simple <- summary(simple_model)
summary_simple
```

```{r}
# Bootstrap the model on the PLS Estimated Model
boot_model <- bootstrap_model(
  seminr_model = simple_model,
  nboot = 5000,
  cores = parallel::detectCores(),
  seed = 123)

# alpha sets the specified level for significance, i.e. 0.05
summary_boot <- summary(boot_model, alpha = 0.05)

# Inspect the bootstrapped structural paths
summary_boot$bootstrapped_paths
```

##Satisfaction with job

```{r}
#what to test
dat <- resp %>% dplyr::select("Q1_1","Q6_1","Q6_2","Q29_4", #local loyalty
                              "Q6_5")
colnames(dat) <- c("loy1","loy2","loy3","loy4","sat")
```

```{r}
table(dat$sat)
```

```{r}
dat <- dat %>% mutate(loy1=dplyr::recode(loy1, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>% 
  mutate(loy2=dplyr::recode(loy2, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy3=dplyr::recode(loy3, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy4=dplyr::recode(loy4, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(sat=dplyr::recode(sat, '1'='1', '2'='1', '3'='2', '4'='3', '9'='-99'))

dat[] <- sapply(dat, as.numeric)
dat <- dat[complete.cases(dat[ , c('loy1', 'loy2', 'loy3', 'loy4')]),]
```

```{r}
dat <- dummy_cols(dat, select_columns = c("loy1","loy2","loy3","loy4","sat"))
head(dat)
```

```{r}
# Create measurement model
simple_mm <- constructs(
  composite("Loyalty", multi_items("loy", 1:4)),
  composite("Unsatisfied", single_item("sat_1")),
  composite("Satisfied", single_item("sat_2")))

# Create structural model
simple_sm <- relationships(
  paths(from = c("Unsatisfied", "Satisfied"), to = "Loyalty"))

# Estimate the model
simple_model <- estimate_pls(data = dat,
 measurement_model = simple_mm,
 structural_model = simple_sm,
 missing = mean_replacement,
 missing_value = "-99")

# Summarize the model results
summary_simple <- summary(simple_model)
summary_simple
```

```{r}
# Bootstrap the model on the PLS Estimated Model
boot_model <- bootstrap_model(
  seminr_model = simple_model,
  nboot = 5000,
  cores = parallel::detectCores(),
  seed = 123)

# alpha sets the specified level for significance, i.e. 0.05
summary_boot <- summary(boot_model, alpha = 0.05)

# Inspect the bootstrapped structural paths
summary_boot$bootstrapped_paths
```

##Commuting

```{r}
#what to test
dat <- resp %>% dplyr::select("Q1_1","Q6_1","Q6_2","Q29_4", #local loyalty
                              "Q53")
colnames(dat) <- c("loy1","loy2","loy3","loy4","com")
```

```{r}
table(dat$com)
```

```{r}
dat <- dat %>% mutate(loy1=dplyr::recode(loy1, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>% 
  mutate(loy2=dplyr::recode(loy2, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy3=dplyr::recode(loy3, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy4=dplyr::recode(loy4, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(com=dplyr::recode(com, '1'='1', '2'='2', '3'='2'))

dat[] <- sapply(dat, as.numeric)
dat <- dat[complete.cases(dat[ , c('loy1', 'loy2', 'loy3', 'loy4')]),]

dat$com[which(is.na(dat$com))] <- -99
```

```{r}
dat <- dummy_cols(dat, select_columns = c("loy1","loy2","loy3","loy4","com"))
head(dat)
colnames(dat)
```

```{r}
# Create measurement model
simple_mm <- constructs(
  composite("Loyalty", multi_items("loy", 1:4)),
  composite("Other settlement", single_item("com_2")))

# Create structural model
simple_sm <- relationships(
  paths(from = c("Other settlement"), to = "Loyalty"))

# Estimate the model
simple_model <- estimate_pls(data = dat,
 measurement_model = simple_mm,
 structural_model = simple_sm,
 missing = mean_replacement,
 missing_value = "-99")

# Summarize the model results
summary_simple <- summary(simple_model)
summary_simple
```

```{r}
# Bootstrap the model on the PLS Estimated Model
boot_model <- bootstrap_model(
  seminr_model = simple_model,
  nboot = 5000,
  cores = parallel::detectCores(),
  seed = 123)

# alpha sets the specified level for significance, i.e. 0.05
summary_boot <- summary(boot_model, alpha = 0.05)

# Inspect the bootstrapped structural paths
summary_boot$bootstrapped_paths
```

##Happiness

```{r}
#what to test
dat <- resp %>% dplyr::select("Q1_1","Q6_1","Q6_2","Q29_4", #local loyalty
                              "Q6_4")
colnames(dat) <- c("loy1","loy2","loy3","loy4","hap")
```

```{r}
table(dat$hap)
```

```{r}
dat <- dat %>% mutate(loy1=dplyr::recode(loy1, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>% 
  mutate(loy2=dplyr::recode(loy2, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy3=dplyr::recode(loy3, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy4=dplyr::recode(loy4, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(hap=dplyr::recode(hap, '1'='1', '2'='1', '3'='2', '4'='3', '9'='-99'))

dat[] <- sapply(dat, as.numeric)
dat <- dat[complete.cases(dat[ , c('loy1', 'loy2', 'loy3', 'loy4')]),]

#dat$com[which(is.na(dat$com))] <- -99
```

```{r}
dat <- dummy_cols(dat, select_columns = c("loy1","loy2","loy3","loy4","hap"))
head(dat)
colnames(dat)
```

```{r}
# Create measurement model
simple_mm <- constructs(
  composite("Loyalty", multi_items("loy", 1:4)),
  composite("Unhappy", single_item("hap_1")),
  composite("Happy", single_item("hap_2")))

# Create structural model
simple_sm <- relationships(
  paths(from = c("Unhappy", "Happy"), to = "Loyalty"))

# Estimate the model
simple_model <- estimate_pls(data = dat,
 measurement_model = simple_mm,
 structural_model = simple_sm,
 missing = mean_replacement,
 missing_value = "-99")

# Summarize the model results
summary_simple <- summary(simple_model)
summary_simple
```

```{r}
# Bootstrap the model on the PLS Estimated Model
boot_model <- bootstrap_model(
  seminr_model = simple_model,
  nboot = 5000,
  cores = parallel::detectCores(),
  seed = 123)

# alpha sets the specified level for significance, i.e. 0.05
summary_boot <- summary(boot_model, alpha = 0.05)

# Inspect the bootstrapped structural paths
summary_boot$bootstrapped_paths
```

##Neighbour

```{r}
#what to test
dat <- resp %>% dplyr::select("Q1_1","Q6_1","Q6_2","Q29_4", #local loyalty
                              "ID2")
colnames(dat) <- c("loy1","loy2","loy3","loy4","nei")
```

```{r}
table(dat$nei)
sum(table(dat$nei))
```

```{r}
dat <- dat %>% mutate(loy1=dplyr::recode(loy1, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>% 
  mutate(loy2=dplyr::recode(loy2, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy3=dplyr::recode(loy3, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy4=dplyr::recode(loy4, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) 

dat[] <- sapply(dat, as.numeric)
dat <- dat[complete.cases(dat[ , c('loy1', 'loy2', 'loy3', 'loy4')]),]

#dat$com[which(is.na(dat$com))] <- -99
```

```{r}
dat <- dummy_cols(dat, select_columns = c("loy1","loy2","loy3","loy4","nei"))
head(dat)
colnames(dat)
```

```{r}
# Create measurement model
simple_mm <- constructs(
  composite("Loyalty", multi_items("loy", 1:4)),
  composite("Dozsa", single_item("nei_2")),
  composite("Ipar", single_item("nei_3")),
  composite("Jeruzsalem", single_item("nei_4")),
  composite("Egyetem", single_item("nei_5")),
  composite("Cholnoky", single_item("nei_6")),
  composite("Jutasi", single_item("nei_7")),
  composite("Kadarta", single_item("nei_8")))

# Create structural model
simple_sm <- relationships(
  paths(from = c("Dozsa", "Ipar", "Jeruzsalem", "Egyetem", "Cholnoky", "Jutasi", "Kadarta"), to = "Loyalty"))

# Estimate the model
simple_model <- estimate_pls(data = dat,
 measurement_model = simple_mm,
 structural_model = simple_sm,
 missing = mean_replacement,
 missing_value = "-99")

# Summarize the model results
summary_simple <- summary(simple_model)
summary_simple
```

```{r}
# Bootstrap the model on the PLS Estimated Model
boot_model <- bootstrap_model(
  seminr_model = simple_model,
  nboot = 5000,
  cores = parallel::detectCores(),
  seed = 123)

# alpha sets the specified level for significance, i.e. 0.05
summary_boot <- summary(boot_model, alpha = 0.05)

# Inspect the bootstrapped structural paths
summary_boot$bootstrapped_paths
```

##Time spent in the city in a year

```{r}
#what to test
dat <- resp %>% dplyr::select("Q1_1","Q6_1","Q6_2","Q29_4", #local loyalty
                              "Q51")
colnames(dat) <- c("loy1","loy2","loy3","loy4","mon")
```

```{r}
table(dat$mon)
sum(table(dat$mon))
```

```{r}
dat <- dat %>% mutate(loy1=dplyr::recode(loy1, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>% 
  mutate(loy2=dplyr::recode(loy2, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy3=dplyr::recode(loy3, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>%
  mutate(loy4=dplyr::recode(loy4, '1'='1', '2'='2', '3'='3', '4'='4', '9'='NA')) %>% 
  mutate(mon=dplyr::recode(mon, '1'='1', '2'='1', '3'='2', '4'='3', '9'='-99'))

dat[] <- sapply(dat, as.numeric)
dat <- dat[complete.cases(dat[ , c('loy1', 'loy2', 'loy3', 'loy4')]),]

#dat$com[which(is.na(dat$com))] <- -99
```

```{r}
dat <- dummy_cols(dat, select_columns = c("loy1","loy2","loy3","loy4","mon"))
head(dat)
colnames(dat)
```

```{r}
# Create measurement model
simple_mm <- constructs(
  composite("Loyalty", multi_items("loy", 1:4)),
  composite("0-6 month", single_item("mon_1")),
  composite("7-11 month", single_item("mon_2")))

# Create structural model
simple_sm <- relationships(
  paths(from = c("Less than 6 month", "7-11 month"), to = "Loyalty"))

# Estimate the model
simple_model <- estimate_pls(data = dat,
 measurement_model = simple_mm,
 structural_model = simple_sm,
 missing = mean_replacement,
 missing_value = "-99")

# Summarize the model results
summary_simple <- summary(simple_model)
summary_simple
```

```{r}
# Bootstrap the model on the PLS Estimated Model
boot_model <- bootstrap_model(
  seminr_model = simple_model,
  nboot = 5000,
  cores = parallel::detectCores(),
  seed = 123)

# alpha sets the specified level for significance, i.e. 0.05
summary_boot <- summary(boot_model, alpha = 0.05)

# Inspect the bootstrapped structural paths
summary_boot$bootstrapped_paths
```

#Chi2 test

```{r}
x <- resp$loy
y <- resp$Q59
table(x,y)
prop.table(table(x,y),1)
prop.table(table(y))
z <- as.data.frame.matrix(table(x,y))
mosaic(table(x,y), shade = TRUE, labeling = TRUE)
```

#LCA (Latent class analysis)

##Setting up

```{r}
f <- cbind(Q1_1,Q6_1,Q6_2,Q29_4) ~ 1
```

```{r}
set.seed(12300)
lc <- poLCA(f, dat_lat, nclass = 4, maxiter = 50000, graphs = FALSE, na.rm = TRUE, nrep =  10, verbose = TRUE)
```

```{r}
table(lc$predclass)
prop.table(table(lc$predclass))
sum(table(lc$predclass))
```

##Plotdata

```{r}
x1 <- as.data.frame(lc$probs[1])[,c(1:4)]
x1$class <- 1:4

x2 <- as.data.frame(lc$probs[2])[,c(1:4)]
x2$class <- 1:4

x3 <- as.data.frame(lc$probs[3])[,c(1:4)]
x3$class <- 1:4

x4 <- as.data.frame(lc$probs[4])[,c(1:4)]
x4$class <- 1:4

longdata <- reshape(data=x1,
                    direction = "long",
                    varying = names(x1)[1:4],
                    v.names = "class",
                    times = names(x1)[1:4])

longdata1 <- reshape(data=x2,
                    direction = "long",
                    varying = names(x2)[1:4],
                    v.names = "class",
                    times = names(x2)[1:4])
longdata <- rbind(longdata,longdata1)

longdata1 <- reshape(data=x3,
                    direction = "long",
                    varying = names(x3)[1:4],
                    v.names = "class",
                    times = names(x3)[1:4])
longdata <- rbind(longdata,longdata1)

longdata1 <- reshape(data=x4,
                    direction = "long",
                    varying = names(x4)[1:4],
                    v.names = "class",
                    times = names(x4)[1:4])
longdata <- rbind(longdata,longdata1)

longdata$variable <- rep(c("Q1_1","Q6_1","Q6_2","Q29_4"), each=16)
longdata$time <- rep(1:4, each=4)

colnames(longdata) <- c("value","time","clust","variable")
```

##Plot

```{r}
fac.variable <- c("Engage",
              "Relevance",
              "Future intention",
              "Rational ties")

names(fac.variable) <- c("Q1_1","Q6_1","Q6_2","Q29_4")
```

```{r}
fac.clust <- c("Loyal\n41%",
                "Least loyal\n9%",
                "Semi-loyal\n13%",
                "Very loyal\n37%")
names(fac.clust) <- c(1,2,3,4)
longdata$clust <- factor(longdata$clust, levels=c(4,1,3,2))
```

```{r}
ggplot(longdata, aes(x=time, y=value)) + geom_bar(color="black", fill="skyblue", 
                                                  stat = "identity") + 
  scale_x_discrete(limits=1:4) +
  labs(x="",y="", title="") +
  facet_grid(clust~variable, 
             labeller=labeller(variable=fac.variable, clust=fac.clust)) +
  theme_bw() +
  theme(strip.text = element_text(family = "Calibri"),
        axis.text = element_text(family = "Calibri"))
```

```{r}
ggsave("LCA_plot.jpg", dpi = 300, height = 5, width = 5.5)
```


#RF (Random Forest)

##Data

```{r}
#dependent: loyalty levels
#independent: all other variables

dat3 <- resp %>% dplyr::select(-one_of("id","Q1_1","Q29_4","Q6_1","Q6_2"))
colnames(dat3)                        
```

```{r}
#some conversions
summary(dat3)
sapply(dat3, class)
colnames(dat3)
colSums(is.na(dat3))
dat3[is.na(dat3)] <- 0

#convert to factor
col.names <- colnames(dat3)
dat3[,col.names] <- lapply(dat3[,col.names], as.factor)
sapply(dat3, class)
```

##Split traint-test

```{r}
set.seed(123)
sample <- sample.split(dat3$loy, SplitRatio = .75)

train <- subset(dat3, sample == TRUE)
test <- subset(dat3, sample == FALSE)

dim(train)
dim(test)

#train <- train %>% dplyr::select(-one_of("Q3x5.11.", "Q3x5.7."))
#test <- test %>% dplyr::select(-one_of("Q3x5.11.", "Q3x5.7."))
```

##Optimize parameters

###ntree

```{r}
model <- randomForest(loy ~ ., data=train, proximity=TRUE, ntree=1000)

model

oob.error.data <- data.frame(
  Trees=rep(1:nrow(model$err.rate), times=5),
  Type=rep(c("OOB", "1", "2", "3", "4"), each=nrow(model$err.rate)),
  Error=c(model$err.rate[,"OOB"], 
    model$err.rate[,"1"], 
    model$err.rate[,"2"],
    model$err.rate[,"3"],
    model$err.rate[,"4"]
))

ggplot(data=oob.error.data, aes(x=Trees, y=Error)) +
  geom_line(aes(color=Type))
```


###mtry

```{r}
# mtry (to control how many variables are considered at each step)
# default value is sqrt(number of variables) for classification

oob.values <- vector(length=15)

set.seed(112200)

for(i in 1:15) {
  temp.model <- randomForest(loy ~ ., data=train, mtry=2*i, ntree=400)
  oob.values[i] <- temp.model$err.rate[nrow(temp.model$err.rate),1]
}

oob.values
## find the minimum error
min(oob.values)
## find the optimal value for mtry...
which(oob.values == min(oob.values))
```


###Fit model

```{r}
set.seed(111)
rf <- randomForest(loy ~ ., data = train, 
                   ntree = 400, mtry = 24, 
                   importance = TRUE)
rf
```

```{r}
colnames(test)
pred <- predict(rf, newdata = test[,-128]) #remove dependent variable

table(test$loy, pred)

sum(diag(table(test$loy, pred)))/sum(table(test$loy, pred))
1-sum(diag(table(test$loy, pred)))/sum(table(test$loy, pred))
```

##Var importance

```{r}
varImpPlot(rf)
```

```{r}
varimp <- as.data.frame(rf$importance)

plotdata <- data.frame(y=rownames(varimp),
                       x=varimp$MeanDecreaseAccuracy)
plotdata <- plotdata %>% arrange(-x) %>% top_n(10)
plotdata$y_names <- c("I want my children to live here",
                      "Attachment to the sub-region",
                      "Living in a neighbourhood",
                      "I try to use local entrepreneur\nto do the work around my house/flat",
                      "Attachment to the region",
                      "I'm satisfy with my work",
                      "I'm a happy person",
                      "I'm satisfied with the historical sites",
                      "Work-life balance",
                      "I'm satisfied with the book stores")
```

```{r}
plotdata %>% mutate(y_names=factor(y_names, levels=y_names)) %>%
ggplot(aes(x=y_names, y=x)) +
  geom_segment( aes(x=y_names, xend=y_names, y=0, yend=x), color="skyblue") +
  geom_point(shape=21, color="skyblue", fill="white", size=2, stroke=2) +
  labs(x="", y="Feature importance", 
       title="") +
  theme_light() +
  coord_flip() +
  scale_x_discrete(limits=rev) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(size=12, family = "Calibri"),
    axis.title = element_text(family = "Calibri"),
    axis.text.x = element_text(family = "Calibri"),
    plot.title = element_text(family = "Calibri", size=12))
```


```{r}
ggsave("varimp.png", dpi=300, height = 4.6, width = 6.3)
```

#Kruskal-Wallis test

```{r}
kruskal.test(Q6_3 ~ loy, data = resp)
kruskal.test(Q1_2 ~ loy, data = resp)
kruskal.test(ID2 ~ loy, data = resp)
kruskal.test(Q29_3 ~ loy, data = resp)
kruskal.test(Q6_5 ~ loy, data = resp)
kruskal.test(Q1_3 ~ loy, data = resp)
kruskal.test(Q6_4 ~ loy, data = resp)
kruskal.test(Q24b_11 ~ loy, data = resp)
kruskal.test(Q6_7 ~ loy, data = resp)
kruskal.test(Q24b_9 ~ loy, data = resp)
```

#ARM (Association rule mining)

Suggestion:
https://www.philippe-fournier-viger.com/spmf/index.php

##Transactions

###Data

```{r}
#create longData and
#modify scale values if needed
longData <- melt(resp, id.vars = "id")

#check
table(longData$value)
```

```{r}
#prepare data to transaction set
longData <- cbind(longData, to.dummy(longData$value, "value"))

longData <- longData %>% dplyr::select(-one_of("value", "value.NA"))
```

###Transaction set

```{r}
#create transaction data frame to arules transactions
transactions <- reshape(longData, idvar = "id", timevar = "variable", direction = "wide")

#which answer categories should be deleted because empty or full
colnames(transactions)[which(colSums(transactions, na.rm=TRUE) %in% 0)]
colnames(transactions)[which(colSums(transactions, na.rm=TRUE) %in% nrow(transactions))]

#empty or full columns should be deleted to avoid bias
transactions <- transactions %>% dplyr::select(which(!colSums(transactions, na.rm=TRUE) %in% c(0,1)))
transactions <- transactions %>% dplyr::select(which(!colSums(transactions, na.rm=TRUE) %in%
                                                nrow(transactions)))

#rename columns to improve readability
#e.g. instead of value.0.Q3x2.7. --> Q3x2.7._0 ("variable name"_"variable value")
#colnames(transactions)
colnames(transactions) <- str_remove(colnames(transactions), "value.")
colnames(transactions) <- paste0(colnames(transactions), "_", 
                                 substr(colnames(transactions), start = 1, stop = 2))
colnames(transactions) <- substr(colnames(transactions), 
                                 start = 3, 
                                 stop = nchar(colnames(transactions))-1)
colnames(transactions)[1] <- "id"
```

```{r}
#create transaction matrix for arules
tr.m <- as.matrix(transactions[,-which(colnames(transactions)=="id")])
tr.m <- as(tr.m, "itemMatrix")
```

##Create length 2 edge list

A --> B association rules (if A then B)

```{r}
edges <- data.frame()

for (i in 1:(ncol(transactions)-1)){
rules <- apriori(data=tr.m,
                 parameter=list(supp = 0.001, conf = 0.001, 
                                minlen = 2, #length 2
                                maxlen = 2, 
                                maxtime = 10000,
                                target="rules"),
                 appearance=list(rhs=colnames(transactions)[i+1]),
                 control = list(verbose=F)
)
intme <- interestMeasure(rules, transactions = tr.m,
                         measure = c("rulePowerFactor", "addedValue", "cosine", "collectiveStrength", "leverage", "kappa", "chiSquared", "phi", "fishersExactTest", "hyperConfidence", "hyperLift", "lambda", "doc", "leastContradiction", "oddsRatio", "yuleY", "RLD", "sebag", "importance", "jaccard", "imbalance", "kulczynski", "conviction", "mutualInformation", "jMeasure"))

quality(rules) <- cbind(quality(rules),intme)
df = DATAFRAME(rules, itemSep = ",", setStart="", setEnd="")
df$LHS <- as.character(df$LHS)
df$RHS <- as.character(df$RHS)
df <- df[df$count!=0,]
edges <- rbind(edges, df)
}

edges$relrisk <- edges$confidence/(edges$confidence-edges$doc) #relative risk not implemented
edges$modcos <- edges$cosine-sqrt(edges$support-edges$leverage) #modified cosine --> own measure
```

```{r}
write.csv2(edges, "edges_length2.csv", row.names = FALSE)
```

##Create length 3 edge list

A&B --> C association rules (If A union/and B, then C)

This rules provided opportunity to analyze the loyalty of people with unsufficient social relationship

```{r}
edges <- data.frame()

rules <- apriori(data=tr.m,
                 parameter=list(supp = 0.001, conf = 0.001, 
                                minlen = 3, #length 2
                                maxlen = 3, 
                                maxtime = 10000,
                                target="rules"),
                 appearance=list(rhs=list("loy_1", "loy_2", "loy_3", "loy_4")),
                 control = list(verbose=F)
)

intme <- interestMeasure(rules, transactions = tr.m,
                         measure = c("rulePowerFactor", "addedValue", "cosine", "collectiveStrength", "leverage", "kappa", "chiSquared", "phi", "fishersExactTest", "hyperConfidence", "hyperLift", "lambda", "doc", "leastContradiction", "oddsRatio", "yuleY", "RLD", "sebag", "importance", "jaccard", "imbalance", "kulczynski", "conviction", "mutualInformation", "jMeasure"))

quality(rules) <- cbind(quality(rules),intme)
df = DATAFRAME(rules, itemSep = ",", setStart="", setEnd="")
df$LHS <- as.character(df$LHS)
df$RHS <- as.character(df$RHS)
df <- df[df$count!=0,]
edges <- rbind(edges, df)


edges$relrisk <- edges$confidence/(edges$confidence-edges$doc) #relative risk not implemented in R
edges$modcos <- edges$cosine-sqrt(edges$support-edges$leverage) #modified cosine --> our measure
```

```{r}
write.csv2(edges, "edges_length3.csv", row.names = FALSE)
```

##Association network

###Edges

```{r}
df1 <- read.csv2(edges_length2.csv, sep = ";")
```

```{r}
#filter correlated rules by several measures
edges1 <- df1[df1$sign==1 & 
                df1$cosine>=0.3 &
                df1$relativeAdded>=0.4 &
                df1$addedValue>=0.05 &
                df1$phi>=0.15 &
                df1$confidence.x>=0.25
              ,]
edges1$LHS <- as.character(edges1$LHS)
edges1$RHS <- as.character(edges1$RHS)
```

```{r}
edge <- data.frame(
  from=edges1[,"LHS"],
  to=edges1[,"RHS"],
  smooth=c(TRUE),
  width=edges1$count/50,
  color="#5080b1",
  stringsAsFactors = FALSE
)
```

###Nodes

```{r}
##node tábla
node <- data.frame(
  id=unique(c(edges1$LHS,edges1$RHS)),
  label=unique(c(edges1$LHS,edges1$RHS)),
  shape="circle",
  shadow=c(FALSE),
  stringsAsFactors = FALSE
)
```

```{r}
node$variable <- gsub("\\..*","",node$label)
node$answer <- gsub(".*\\.","",node$label)
node$answer <- as.numeric(node$answer)
```

```{r}
#color of nodes in general
node$color.border[node$answer==4] <- "#77933c" #green
node$color.border[node$answer==3] <- "#c3d69b" #light green
node$color.border[node$answer==2] <- "#e6b9b8" #right red
node$color.border[node$answer==1] <- "#953735" #red
```

```{r}
#color of exceptional nodes (some variable measured with reversed scale or dummy)
node$color.border[node$answer==1 & node$variable=="Q33a"] <- "#953735" #red 
node$color.border[node$answer==2 & node$variable=="Q33a"] <- "#77933c" #green

node$color.border[node$answer==1 & node$variable=="Q31"] <- "#953735" #red
node$color.border[node$answer==2 & node$variable=="Q31"] <- "#77933c" #green

node$color.border[which(node$answer==1 & startsWith(node$variable, "Q32"))] <- "#953735" #red
node$color.border[which(node$answer==2 & startsWith(node$variable, "Q32"))] <- "#77933c" #green

#dependent variables
node$color.border[which(node$variable %in% c("loy","dev","bet","pri"))] <- "#80ffff" #turquoise

#demography
node$color.border[which(node$variable %in% c("ID","gen","Q50","Q57","Q59"))] <- "#3399ff" #dark blue

```

```{r}
node$color.background <- "white"
node$color.highlight <- "yellow"
```

###Plot

```{r}
nw <- 
  visNetwork(node, edge, height = "1000px", width = "100%") %>% 
  visOptions(highlightNearest = TRUE, selectedBy = "label") %>% 
  visInteraction(dragNodes = TRUE, 
                 dragView = TRUE, 
                 zoomView = TRUE,
                 hideEdgesOnDrag = FALSE) %>%
  visNodes(physics=FALSE, size=16, borderWidth = 1, font=list(face="calibri")) %>%
  visEdges(arrows=list(to=list(enabled=T, scaleFactor=1))) %>% 
  visIgraphLayout(#layout = "layout_in_circle",
                  type="full")

visSave(nw, file="nw_associations_20200602_v1.html")
```

##Association subnetwork

This code sequence draw a piece of the whole network.

###Edges

```{r}
#Filter the questionable LHS-RHS

#social relational (could get help in problems) & loyalty level variables
df1_sub <- df1[startsWith(df1$LHS,"Q16c") | startsWith(df1$LHS,"Q16d") | 
                 startsWith(df1$LHS, "loy"),]
df1_sub <- df1_sub[startsWith(df1_sub$RHS, "loy") | startsWith(df1_sub$RHS,"Q16c") | startsWith(df1_sub$RHS,"Q16d") ,]

#subset about get help from friends, neighbours, colleagues
df1_sub <- df1_sub[which(sub("\\..*", "", df1_sub$LHS) %>% sub(".*_", "", .) %>% as.numeric() %in% c(1,2,3,4)),] 

df1_sub <- df1_sub[c(which(sub("\\..*", "", df1_sub$RHS) %>% sub(".*_", "", .) %>% as.numeric() %in% c(1,2,3,4)), which(startsWith(df1_sub$RHS, "loy"))),] 

#significant associations
df1_sub <- df1_sub[df1_sub$sign==1,]

#more correlated rules
df1_sub <- df1_sub[df1_sub$cosine>=0.18 & 
                     df1_sub$addedValue>=0.03 & 
                     df1_sub$relativeAdded>=0.08,]
```

```{r}
edge <- data.frame(
  from=df1_sub[,"LHS"],
  to=df1_sub[,"RHS"],
  smooth=c(TRUE),
  width=df1_sub$count/50,
  color="#95b3d0", #"#5080b1",
  stringsAsFactors = FALSE
)
```

###Nodes

```{r}
node <- data.frame(
  id=unique(c(df1_sub$LHS,df1_sub$RHS)),
  label=unique(c(df1_sub$LHS,df1_sub$RHS)),
  shape="circle",
  shadow=c(FALSE),
  stringsAsFactors = FALSE
)
```

```{r}
node$variable <- gsub("\\..*","",node$label)
node$answer <- gsub(".*\\.","",node$label)
node$answer <- as.numeric(node$answer)
```

```{r}
#color of nodes in general
node$color.border[node$answer==4] <- "#77933c" #green
node$color.border[node$answer==3] <- "#c3d69b" #light green
node$color.border[node$answer==2] <- "#e6b9b8" #light red
node$color.border[node$answer==1] <- "#953735" #red
```

```{r}
#color of exceptional nodes (some variable measured with reversed scale or dummy)
node$color.border[node$answer==1 & node$variable=="Q33a"] <- "#953735" #red 
node$color.border[node$answer==2 & node$variable=="Q33a"] <- "#77933c" #green

node$color.border[node$answer==1 & node$variable=="Q31"] <- "#953735" #red
node$color.border[node$answer==2 & node$variable=="Q31"] <- "#77933c" #green

node$color.border[which(node$answer==1 & startsWith(node$variable, "Q32"))] <- "#953735" #red
node$color.border[which(node$answer==2 & startsWith(node$variable, "Q32"))] <- "#77933c" #green

#explained variables
node$color.border[which(node$variable %in% c("loy","dev","bet","pri"))] <- "#80ffff" #turquoise

#demography
node$color.border[which(node$variable %in% c("ID","gen","Q50","Q57","Q59"))] <- "#3399ff" #dark blue

```

```{r}
node$color.background <- "white"
node$color.highlight <- "yellow"
```

```{r}
node[node$variable %in% c("loy","dev","bet","pri"),"shape"] <- "triangle"
```

```{r}
node[which(startsWith(node$label,"Q16c")), "shape"] <- "box"
```

###Plot

```{r}
nw <- 
  visNetwork(node, edge, height = "1000px", width = "100%") %>% 
  visOptions(highlightNearest = TRUE, selectedBy = "label") %>% 
  visInteraction(dragNodes = TRUE, 
                 dragView = TRUE, 
                 zoomView = TRUE,
                 hideEdgesOnDrag = FALSE) %>%
  visNodes(physics=FALSE, size=16, borderWidth = 1, font=list(face="calibri")) %>%
  visEdges(arrows=list(to=list(enabled=T, scaleFactor=1))) %>% 
  visIgraphLayout(#layout = "layout_in_circle",
                  type="full")

visSave(nw, file="nw_associations_loy_20200602_v2.html")
```






